---
output: 
  revealjs::revealjs_presentation:
    theme: white
    highlight: pygments
    css: custom.css
    center: true
    transition: none
    reveal_options:
      slideNumber: 'c/t'
      progress: true
    fig_width: 8.0
    fig_height: 4.0
    fig_caption: true
    self_contained: true
    # reveal_plugins: ["notes", "chalkboard"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(scales)
library(kableExtra)
library(dplyr)
library(captioner)
library(stringr)
source("code/cc_quality_assimilation_flow_presentation.R")
```


<!-- FontAwesome -->
<link rel="stylesheet" href="https://use.fontawesome.com/releases/v5.11.2/css/all.css" crossorigin="anonymous">

----

## Labor-Market Quality <br> of Foreign-Born Workers <br> 2007-2017

### Rubén Hernández Murillo

FDIC — May 13, 2022

(_Federal Reserve Bank of Cleveland **Economic Commentary**_, May 2020.)

----

#### Goals

+ The foreign-born are (**_very_**) important for the US economy
+ Data from the ACS on _labor-market quality_ between 2007 and 2017
+ Various measures of labor-market quality **improved** over the period
  - Notable increases among Mexican-born workers
+ Why?
  + Shift in the composition of the foreign-born
    - Increase in immigration from Asia
    - Precipitous decline in immigration from Mexico (*why?*)

----

### The foreign-born are very important

+ Most of the foreign-born population are working-age individuals
  - The foreign-born represent almost 20% of the US workforce
+ The foreign-born population grew **2.5x** to **3.3x** **faster** than the US-born
  - Overall US population grew 8.2%:
    - 2.2% was growth in the _overall_ foreign-born population
    - 6.1% was growth in the _overall_ US-born population
  - Working-age US population grew 5.7%:
    - 2.2% was growth in foreign-born **workers**
    - 3.5% was growth in US-born **workers**

----

### Measures of labor-market quality

<ul class="fa-ul">
<li><span class = "fa-li"><i class="fas fa-book-reader"></i></span> Educational attainment</li>
<li><span class = "fa-li"><i class="fas fa-comment-dots"></i></span>English proficiency</li>
<li><span class = "fa-li"><i class="fas fa-money-bill-wave"></i></span>Earnings potential</li>
</ul>

----


### Educational attainment by area of origin

```{r, echo=FALSE}
plot(plot.edu)
```

----

### English proficiency by area of origin

```{r echo=FALSE}
plot(plot.eng)
```

----

### Median annual earnings relative to <br> US-born workers by area of origin

```{r echo=FALSE}
plot(plot.earn)
```

----

### Distribution of the recent foreign-born <br> by area of origin

```{r echo=FALSE}
plot(p.waob.shares.ff)
```

----

### Border enforcement

```{r echo=FALSE, fig.height=5.5}
plot(grid.enforcement)
```


----

### Conclusion

- Measures of labor-market quality of foreign-born workers <br> **improved** during 2007-2017
- Notable gains seem concentrated among immigrants from **Mexico**
- The apparent increase in quality results from <br> a **shift in the distribution of immigrants**:
  + Increased immigration from Asia
  + A precipitous decline in immigration from Mexico
- Factors behind the decline in immigration from Mexico:
  - Increase in **border enforcement** over past two decades
  - (Decrease in US demand for Mexican workers after 2008/9 recession)